### 2015-06-08 LOI/COI/COI-lofstead.tex
added all of the kinds of relationship data

  \item Hasan Abassi, Oak Ridge National Laboratory, collaborator
  \item Andrea Arpaci-Dusseah, UW Madison, co-author
  \item Remzi Arpaci-Dusseau, UW Madison, co-author
  \item Roselyne Barreto, University of Chicago, co-author
  \item John Bent, EMC Corporation, co-author
  \item Michael Booth, Oracle, co-author
  \item Franck Cappello, Argonne National Laboratory, collaborator
  \item Jacqueline Chen, Sandia National Laboratories, co-author
  \item Julian Cummings, California Institute of Technology, co-author
  \item Jai Dayal, Georgia Institute of Technologies, collaborator
  \item Ciprian Docan, Rutgers, co-author
  \item Stephane Ethier, Princenton Univeristy, co-author
  \item Ada Gavrilovska, Georgia Institute of Technology, co-author
  \item Garth Gibson, Carnegie Mellon University, co-author
  \item Ray Grout, National Renewal Energy Laboratory, co-author
  \item Gregory Jean-Baptiste, Florida International University, co-author
  \item Ivo Jimenez, University of California Santa Cruz, co-author
  \item Scott Klasky, Oak Ridge National Laboratory, co-author
  \item Todd Kordenbrock, Hewlett-Packard, collaborator
  \item Seung-Hoe Ku, New York University, co-author
  \item Qing Liu, Oak Ridge National Laboratory, co-author
  \item Mike Lang, Los Alamos National Laboratory, collaborator
  \item Jack Lange, University of Pittsburgh, collaborator
  \item Jeremy Logan, Oak Ridge National Laboratory, co-author
  \item Xiaosong Ma, North Carolina State University, co-author
  \item Carlos Maltzahn, University of California Santa Cruz, collaborator
  \item Kathryn Mohror, Lawrence Livermore National Laboratory, collaborator
  \item Adam Moody, Lawrence Livermore National Laboratory, collaborator
  \item Ron Oldfield, Sandia National Laboratoriey, collaborator
  \item Manish Parashar, Rutgers, co-author
  \item Tom Peterka, Argonne National Laboratory, collaborator
  \item Norbert Podhorzski, Oak Ridge National Laboratory, co-author
  \item Milo Polte, WibiData, co-author
  \item Charles Reiss, University of California Berkeley, co-author
  \item Karsten Schwan, Georgia Institute of Technology, PhD advisor
  \item Arie Shoshani, Lawrence Berkeley National Laboratory, co-author
  \item Alexander Sim, Lawrence Berkeley National Laboratory, co-author
  \item Gregory Sjaardema, Sandia National Laboratories, co-author
  \item Sagar Thapaliya, University of Alabama, Birmingham, collaborator
  \item Yuan Tian, Oak Ridge National Laboratory, co-author
  \item Yandong Wang, Auburn University, co-author
  \item Meghan Wingate McClelland, Xyratex, co-author
  \item Weikuan Yu, Auburn University, co-author
  \item Fang Zheng, IBM, co-author