### 2015-06-17 narrative/deliverables.tex
added some outline material and instructions in the introduction

\section*{Deliverables}

\begin{itemize}
\item plug-in architecture for Sirocco
\item storage APIs to manipulate storing and retrieving data to specific tiers
\item metadata management for data stored in multiple tiers using different compression
\item metadata management for deeper data knowledge
\item middleware hooks to manage new storage APIs
\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
\end{itemize}