### 2015-06-17 narrative/management.tex

\section*{Management}

\begin{itemize}

\item We need to manage things

\end{itemize}