### 2015-06-22 other-support/lofstead-cp.tex

\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{cite}
\usepackage{url}
\usepackage{times}
\usepackage{graphicx}
\usepackage[sectionbib]{bibunits}
\pagestyle{empty}

% Set page geometry
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

\begin{center}
{\Large Current and Pending Support for Gerald F. Lofstead}
\end{center}

Dr. Lofstead is a senior member of the technical staff in the Scalable
System Software group at Sandia National Laboratories (SNL). The following
table identifies current and pending support. This funding provides for effort,
travel, and resources. The effort levels will be adjusted if pending proposals
are funded, and as current projects finish.

\section*{Current Support}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Scalable I/O Services\\
Source of Support           \>: \> NNSA Advanced Simulation and Computing Program\\
Award Amount                \>: \> \$1000K/year to SNL\\
Period Covered              \>: \> 10/2014 -- 9/2014\\
Investigator Commitment     \>: \> 0.2 FTE
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= ATDM Data Management Services\\
Source of Support           \>: \> NNSA Advanced Simulation and Computing Program\\
Award Amount                \>: \> \$9800K/year to SNL\\
Period Covered              \>: \> 10/2014 -- 9/2018\\
Investigator Commitment     \>: \> 0.25 FTE
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \=  HOBBES: OS and Runtime Support For Application Composition\\
Source of Support           \>: \> 2013 DOE Exascale Operating Systems and Runtimes\\
Award Amount                \>: \> \$340K/year to SNL\\
Period Covered              \>: \> 8/2013 -- 7/2016\\
Investigator Commitment     \>: \> 0.35 FTE
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= High-Performance Decoupling of Tightly-Coupled Data Flows\\
Source of Support           \>: \> DOE Scientific Data Management, Analysis and Visualization at Extreme Scale 2\\
Award Amount                \>: \> \$200K/year to SNL\\
Period Covered              \>: \> 8/2014 -- 7/2017\\
Investigator Commitment     \>: \> 0.2 FTE
\end{tabbing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Pending Proposals (in addition to this)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Pending Proposals (in addition to this)}

None

%\begin{tabbing}
%Project Title \hspace{5.5em}\=: \= Code Coupling for Scalability and Robust Data Analytics\\
%Source of Support           \>: \> Sandia National Laboratories LDRD \\
%Award Amount                \>: \> \$600K/year to SNL\\
%Period Covered              \>: \> 8/2014 -- 7/2017\\
%Investigator Commitment     \>: \> 0.5 FTE
%\end{tabbing}

\end{document}