### 2015-06-26 narrative/data-reorg.tex

\subsubsection{Data Reorganization Techniques}

Mark, Scott