### 2015-07-01 narrative/eviction-metrics.tex
modified:   narrative/eviction-metrics.tex 	modified:   proposal.tex

\subsubsection{Eviction Metrics}

% The title might have to change...

There is a spectrum of approaches on how to leverage a memory and storage hierarchy: On one end of the spectrum is to copy or move data from one tier to another, the closer the tier to the hierarchy's top, the faster the access to the data it contains. On the other end of the spectrum is to use higher tiers exclusively for auxiliary data that to some extent represents the actual data on the lowest tier. Examples of auxiliary data are views, indices, lossy compressions, lower resolution data, and summary data (sometimes also known as metadata). There are a number of drivers that push