### 2015-07-02 narrative/qos-storage.tex
modified:   ../bib/maltzahn.bib 	modified:   infrastructure.tex 	deleted:    policies.tex 	modified:   proposed-work.tex 	modified:   qos-storage.tex 	deleted:    time-estimation.tex

\subsubsection{QoS}

\paragraph{Background:} One of the key differentiating features this project offers is driving storage
decisions through quality of service (QoS) \emph{reservations} rather than continuing the current model of allowing
all application to compete, without restriction, for part or all of the storage
resources. This traditional approach leads to intereference
effects~\cite{lofstead:2010:adaptive} that can greatly impact IO performance
predictability. Further, with the introduction of additional tiers in the
storage stack, the performance variability will increase with greater
competiion for limited high performance resources. Unfettered, this competition
will yield lower performance for all users than if applications attempt to reserve performance, thereby enabling the storage system to reason over its performance in terms of reservations and actual performance, accepting reservations that are within resource availability, rejecting those that would lead to overload.

It is important to note that we use performance reservations to reduce interference and to avoid overload conditions. If the system uses reservations intelligently, its performance is predictable and efficient because resources are not wasted on interference. However, in cases where an appication arrives that needs to run more urgently than any of the other already running applications, administrators can suspend less urgent applications which will save their state and relinquish their performance reservations. Upon restart, these applications make new performance reservations.

There are two fundamental causes that can impede the I/O performance in an extreme-scale SSIO system. One is indirection, where layers of storage tiers are employed either for performance and/or for scalability reasons.  The direct consequence is that there could be multiple traversing paths from application end to the rest place of data, and more often than not, the I/O paths are not under control under any single authority. The other cause is the shared use of resources. The best effort I/O request/response nature and lack of QoS mechanisms imply that there is little guarantee in terms of expected performance.  Both indirection and shared use of resource contribute to a high probability of imbalanced use of resources thus the occurrence of congestion and degraded performance. Traditionally, there is a disconnection between storage infrastructure and rest of system software and applications: the infrastructure details are not exposed anywhere at all. One argument favoring this is to ensure a platform agnostic design.

As an example to demonstrate that this disconnection will hurt system performance: We launch 4096 processes with each process doing a single file I/O operation against half of the Spider II file system. The traces of those files are analyzed to examine the utilization distribution of different components. Figure (a), (b) and (c) shows the resource usage distribution for OSTs, OSSes, and LNETs, respectively. We observe that there exists a significant variation in usage across components of any given type (e.g., OST, OSS or LNET). For example, some OSTs are used more than 10 times while some others are never used (corresponding to zero frequency count). Similarly, OSSes and LNETs show significant imbalance in usage under the default placement strategy. Consequently, imbalanced resource utilization increases the contention at certain components more than others.

Given these insights, we advocate the idea that the infrastructure knowledge should find a way to relay to the upper layer for better and more effective use of resources.   We think this is more pertinent and critical given the recent development of multi-tier storage and storage component heterogeneity.  There needs to be a way for application/middleware layer to gain more exposure of storage system for more intelligent processing logic.  One prime example of such exposed knowledge can be request/response time. To most applications, this is a black box. Profiling it at the upper layer is neither efficient nor effective, as it doesn't reflect cross-layer characteristics. However, Most of storage layer does keep a detailed profiling of such information. We therefore envision and propose a histogram-based request/response profiling API that application and middleware layer can leverage and make more informative decisions.

\begin{figure}[tbh]
  \centering
  \includegraphics[width=\columnwidth]{graphics/infrastructure.pdf}\vspace{-1.2in}
  \caption{Resource usage distribution for OST(a), OSS(b) and LNETs(c). }
\end{figure}

% This is about finding the relationshiop of workload-dependent to workload-independent metrics. -- Carlos
 %are driven from multiple different approaches. First, supporting
%capability runs is a priority for capability machines. We propose to address
%capability run performance by incorporating monitoring for prepartory runs at
%smaller scale to profile the application output characteristics both from a
%writing velocity and volume, but also for example read patterns for the data
%analytics required to generate scientific insights from the raw data.

% Not sure how to reuse this ... -- Carlos
%By
%discovering approximate data proportions that will likely be the key subsets
%targeted by the simulation run, we can generate a policy to preserve a certain
%storage portion for high fidelity data storagae and use slower or perhaps lower
%fidelity or compressed data storage for less interesting data portions. We will
%support spreading data appropriately across the storage hierarchy. In
%particular, we will investigate how to place data across different layers to
%meet the performance requirements for output while maintaining system
%availability for other applications and offering the best possible performance
%for the data analytics that will ultimately process this data.

% Not sure how to reuse this... -- Carlos
%Second, once data has been placed for future performance requirements, policies
%must be able to address maintaining data on a priority basis in that location
%to deliver on the placement optimization generated by the capability run
%profilier.

% This is a matter of metering, not performance reservations. -- Carlos
%Third, most future NVM devices have limited write endurance prompting
%management to ensure fair use by all machine users. Technologies like
%NAND-flash and Phase Change Memory have limited write endurance. By
%incorporating policies about the proportion of write endurance a compute
%allocation is entitled to, data placement decisions can be made to ensure fair
%resource usage. While the offending application will suffer worse IO
%performance, other applications that more carefully address their data
%intensity on these limited devices will achieve higher overall performnce. Only
%by instituting such policies can we encourage application users to adopt
%technology to manage the limited resources.

%We propose to investigate both offering a policy mechanism and how to implement
%these kinds of policies in a way that addresses offering the shortest
%end-to-end time for scientific insights.

``Quality of Service'' (QoS) refers to the
properties of the performance qualtiy of a particular service, in this case the storage system.
Performance quality can be expressed in either relative or absolute
terms. Examples of relative performance terms are ``fair'',
``proportional'', or ``priority/class-based'' (``the higher the
priority the better the service'' or ``1st class is better than
2nd''). The key advantage of relative performance terms is that
they are easy to implement. The key disadvantage is the difficulty
of making any guarantees based on those implementations --
even relative guarantees are mired by the well-known effect of
priority inversion~\cite{lampson:cacm80}. Examples of absolute
performance terms are ``rate-based'', ``soft real-time'', and
``hard real-time''. The key advantage of absolute performance
terms are that they enable the implementation of strong
guarantees: absolute terms enable agreements between application tasks and storage that we call \emph{reservations} and that do not change meaning depending
on the storage system's workload. The key challenge of absolute
performance terms is that their implementation is non-trivial,
especially in large-scale systems.

\paragraph{Approach:} The goals of this proposal require applications
to demand guarantees in terms of absolute performance qualities.
The first key enabler for any service to make absolute guarantees
is \emph{performance isolation}, i.e. the ability to shield the
performance impact of one task from another. The most straight-forward
but least flexible strategy is to simply cap the latency of each
task, irregardless of its completeness (see for
example~\cite{decandia:sosp07}). This strategy is applicable when
incomplete tasks still have value and missing results are either
not important or can be easily retrieved in subsequent tasks. A
more generally applicable strategy is to control the admission of
tasks to the system: \emph{admission control} maintains a
\emph{utilization model} that can predict the utilization of system
resources given a new task, estimates the current utilization of
the system, and decides whether the task can be admitted without
overloading the system. Thus, admission control avoids system
overload conditions that lead to hard-to-model chaotic behaviors.
Consequently, utilization models can frequently be approximated by
automatically calibrated linear models (see for
example~\cite{skourtis:hpdc12}).

Another essential component of performance isolation is the
\emph{charging model} which determines which task(s) should be
charged how much for a request. Performance isolation is implemented
by the charging model by ensuring that performance of a task only
depends on its reservation and its workload behavior (but not on
other workloads in the system). Each task has a budget based on its
reservation and spends it based on the cost of each request determined
by the charging model. A charging model determines the cost based
on the utilization model and the interactions a request can have
with other requests (e.g. a random request to spinning media
interrupting a stream of sequential requests).

Performance isolation (including its utilization and charging models)
critically depends on \emph{workload-independent metrics}. Even
though throughput and latency are frequently the most meaningful
metrics for an application, they are not workload-independent and
therefore are inadequate for measuring the utilization of a device
across all workloads, unless one accepts worst-case estimates that
can lead to gross underutilization of resources (e.g. the throughput
of random and sequential workloads in spinning media differ by
orders of magnitudes). A common workload-independent metric is
\emph{time utilization}, i.e. the amount of time a resource is
utilized in a given time interval. Time utilization has the advantage
of having an always defined maximum and therefore makes resources
fully reservable. Even though a workload-independent metric might
at first not appear meaningful for an application, given a particular
workload an application can discover the relationship between its
workload-dependent metric and the workload-independent metric.
Because of performance isolation, the application has to discover
this relationship only once.

\subsubsection{QoS}

\paragraph{Related Work:}

\paragraph{Challenges:} The deep and heterogenous memory and storage
hierarchy we are assuming for this proposal complicates the
relationship between workload-dependent and workload-independent
metrics: the performance of a task can significantly differ depending
on what levels of the hierarchy are involved. Furthermore, an
important goal of the proposed project is to enable applications
to reason about the trade-off between resolution and latency, adding
yet another dimension to how tasks are mapped to resources.

\begin{itemize}

\item A key challenge of admission control is scalability: the
decision of whether to admit a task could potentially depend on
global knowledge of the current utilization of every single resource
in the system. Scalability therefore depends on whether admission
control is able to accurately map a new task to a relatively small
set of resources which can quickly provide up-to-date utilization.
One approach might involve pseudo-random mapping that also load-balances
as a side-effect.

\item To offer latency/latency trade-offs, the system must be able
to quickly generate a number of data production plans, involving
different parts of the hierarchy and different data resolutions.
It will then use the utilization model to estimate the latency for
each production plan and the resolutions they can provide. Here
again, pseudo-random selection could reduce the number of resources
that would be involved, thereby increasing scalability.

\item The complexity of a utilization model involving the entire
storage hieararchy with 100,000s of devices is potentially daunting.
However, the utilization model can be simplified by modeling classes
of devices as well as classes of requests. In particular, each
device could restrict access to its content via a set of well-defined
methods with known, absolute performance properties.

\item Flash devices become inherently unpredictable when reads and
writes are mixed on the same device because of garbage collection.
While write latency can be easily hidden using asynchronous I/O,
hiding read latency is more difficult. By separating reads and
writes for each device, read latency becomes predictable. The
challenge is to minimize duplication of data, at least on fast
layers and leverage redundancy across layers.

\end{itemize}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End: