### 2015-07-06 narrative/evaluation.tex

\section{Evaluation}
\label{sec:evaluation}

In order to determine if our guiding principles are being met, and we can
optimize the system from these guiding principles, we will evaluate our work
using the Table~\ref{table:eval}.

To determine the effectiveness of the components of our proposed work we will
partition the problem along the lines of our two guiding principles. 
We will evaluate how the knowledge centric system design we employ can
provide a more predictable level of performance and data quality.

%I dont think this table is right. Knowledge is our independent variable
%and predictability is our dependent variable
\begin{tabular}{ | l | l | l | }

  \hline
  {\bf Principle/Metric}                   & {\bf User perspective metric}   & {\bf System perspective metric }\\ \hline
  {\bf Knowledge-centric}                  & More knowledge = faster access  &  More knowledge = better \\
  {\bf system}                             & to higher quality data          &  resource utilization across \\
                                           &                                 & the entire hierarchy \\ \hline

  {\bf Predictable Performance} & More predictability =                     &  Accuracy of the time bounds \\
  {\bf Accuracy}                & better quality of data                    &  across multiple layers of the\\
                                           &                                           &  storage hierarchy wrote/read \\ 
                                           &                                           &  in time constrained\\
                                           &                                           &  situations \\ \hline

\end{tabular}

In addition to benchmarking the various S2E2 methods and tools in isolation,
we will work with our science partners at Princeton University (J. Tromp),
PPPL (C. S. Chang), and U. C. Irvine (Z. Lin), along with the OLCF and SNL,
to evaluate the leading edge applications along with the system performance.
We will use the systems at both SNL and OLCF, which have multiple layers of
storage. For example, the sith cluster at the OLCF
(https://www.olcf.ornl.gov/computing-resources/sith/) has its own Lustre
storage system, SSDs on each node, and is connected to the centerwide Lustre
file system. We will perform test in both the system level (on sith) and in
the user level (on titan), and evaluate our results using our application
knowledge.

Our evaluation will include how accurately we can set the bounds on
time estimates for applications. We will measure these estimates when data
is on just one level of the storage hierarchy, along with the time estimates
when data chunks are split. We will evaluate how fast data can be discovered
when it is split between the hierarchy, allowing us to better understand how
many pieces data can be split into before it takes too long to access most
of the data. We will also evaluate our techniques to examine how well we can
migrate data and update our server to determine if data might get
out-of-sync between data chunks. We will need to clearly understand how well
this system works on current systems and we will need to investigate this on
the next generation OLCF system (Summitt) which we will have early access,
along with all of the applications we will be investigating.

Another area of investigation will be in how accurate data can be read in
when there are time constraints due to users wanting to read in many time
slices in a given amount of time. For example, in J. Tromp's visualization,
we may contain hundreds of time slices, but when we visualize this data we
would like very fast access to the ``first look' of the data. The question
is whether the data will be highly representative of the overall data. Can
we make accurate visualization from the less accurate data? We will work
with our application partners to access this, and we will determine the
accuracy in time-constrained test vs. test which will not place any bounds
on the time to retrieve the data.

We will also want to assess how well our data migration and eviction
policies work with user-knowledge. Will the overall system be able to
utilize more of the storage hierarchy by understanding these intentions and
creating autonomic rules to self-optimize? Furthermore we will want to see
if users can give us more accurate policies which can then be used for
better system resource utilization. If we have more knowledge we would like
to be able to illustrate that this will translate into higher levels of
system utilization.

One of the risks we face in our design is the unwillingness of users to add
more knowledge to the system. To better understand the burden on the user we
will attempt to quantify the additional burden on the users in relation to
the additional knowledge required for optimization. 
 We realize that initially it will be a large burden for users to
place initial knowledge into the system, but we will evaluate the results as
users increase the knowledge placed into the system with the overall burden.
Our hypothesis from working with many LCF users is that once they get over
the initial burden of placing this knowledge the burden decreases as more
knowledge is placed into the system, and thus, they will be more inclined to
increase the knowledge into the system as the benefits increase.

We will also need to evaluate how more knowledge into the system will lead
into more performance they get from placing the most important pieces at the
higher layers of the hierarchy. We will evaluate this by increasing and
decreasing this knowledge. For example, if the user doesn't want their data
broken into priority chunks, then the data will be at the mercy of what the
system does with no knowledge. If the user then gives us information that
allows us to use ``generic'' rouitnes based on frequency, or accuracy, then
we can use that knowledge to decompose the objects into chunks which can be
placed. We can evaluate these against users who will give us more knowledge
which allows us to perform application level binning of data (for example
for C.S. Chang), or for level of features (for J. Tromp).

% How will we quantify the quality of data? We will look at errors after
% analytics.

% How well does the system utilize all of its resources. We feel the more
% knowledge means the better the system can utilize all of the resources.

%predictability ... are the errors bounded,

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End: