### 2015-07-07 appendix-data-management-plan.tex
DM plan

\section*{Appendix 6: Data Management plan}

\vspace{1ex}
\noindent
The relevant data for the proposed work will be from the software we develop from our team, the documentation for our software, the performance data we use in our publications,
and the requirement documents we get from any of the users who we are working with.

Our team is  will be sharing these data artifacts with the broader community. ORNL will setup a dedicated website for this project which will have all relevant information, along with
a link to all of the new software that will be developed using an open source license.  The source code will be maintained using a versioning system and repository, Github, that is widely used and that we expect will also be well known for the next 10 years.

The documentation will be made available online on the project web site. The performance data will be distributed through publications and technical reports. Publications and technical reports will be made available on the project web site as well. In addition, journal, conference, and workshop papers are also typically archived by publishers (e.g. ACM, IEEE, or Springer). Similarly, Oak Ridge and Sandia National Laboratories maintain archives of all the technical reports. The project web site will maintain pointers to all publications and reports, along with all data that was used in the publications.