### 2015-07-07 narrative/deliverables.tex

\section{Deliverables}
The following are the per year project deliverables:

\textbf{Year 1}
\begin{itemize}
\item Develop new techniques for data description that allow users to 
describe the data utility based on their expectations. (ORNL)

\item Assemble and evaluate general techniques for refactoring of data
that will be embedded in the system and can be selected by users or applications. (Brown/ORNL)
\item Explore the use of application hints and utility of data to guide the initial placement of data. 
(Rutgers/ORNL)
\item Investigate the trade-off between ``filing'' and ``piling'' data. (UCSC/Sandia)
\item Demonstrate a metadata service capable of serving both POSIX clients and our clients,
but without consideration for authentication, authorization, or data migration. (Sandia/UCSC)
\end{itemize}

\textbf{Year 2}
\begin{itemize}
\item Develop a new type of querying system that allows an application to ask the storage system
about completion timing information. (ORNL)
\item Research autonomic data management strategies that
can evaluate utility/cost tradeoff, and appropriately place/move data objects at runtime.
\item Demonstrate a time bounded search approach for finding data within the storage system
to identify data and the current location. This will update the cached information in the metadata
service to short-circuit future requests for the data presuming that it does not migrate again soon. 
\item Extend Sirocco to enforce data-centric metrics that help guide the QoS decisions.
(Sandia/UCSC)
\end{itemize}

\textbf{Year 3}
\begin{itemize}
\item Work with DOE applicationexsiting ADIOS-enabled applications, framework using many of the applications to which we have access, including XGC1, GTC, and SPECFM3D
%\item storage APIs to manipulate storing and retrieving data to specific tiers
%\item metadata management for data stored in multiple tiers using different compression
%\item metadata management for deeper data knowledge
%\item middleware hooks to manage new storage APIs
%\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
\end{itemize}