### 2015-07-07 other-support/curry.tex

\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{cite}
\usepackage{url}
\usepackage{times}
\usepackage{graphicx}
\usepackage[sectionbib]{bibunits}
\pagestyle{empty}

% Set page geometry
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

\begin{center}
{\Large Current and Pending Support for Matthew L. Curry}
\end{center}

\section*{Current Support}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Scalable Data Management\\
Principal Investigator      \>: \> Gerald F. Lofstead\\
Source of Support           \>: \> NNSA/Advanced Simulation and Computing\\
Period Covered              \>: \> 10/2014-9/2015\\
Commitment                  \>: \> 0.05 FTE\\
Description                 \>: \> Research and development activities to improve NNSA application performance.\\
Overlap                     \>: \> None\\
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Tajique\\
Principal Investigator      \>: \> Anthony Vestal\\
Source of Support           \>: \> OGA\\
Period Covered              \>: \> 10/2014-9/2015\\
Commitment                  \>: \> 0.1 FTE\\
Description                 \>: \> Characterize distributed file system behavior\\
Overlap                     \>: \> None\\
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Sirocco\\
Principal Investigator      \>: \> Lee Ward\\
Source of Support           \>: \> NNSA/Advanced Simulation and Computing\\
Period Covered              \>: \> 10/2014-9/2015\\
Commitment                  \>: \> 0.85 FTE\\
Description                 \>: \> Development activities for Sirocco, a novel storage systemfor Exascale computers.\\
Overlap                     \>: \> None\\
\end{tabbing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Pending Proposals (in addition to this)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Pending Proposals (in addition to this)}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Managing Node-Based NVMs for HPC Clusters\\
Principal Investigator      \>: \> Ethan Miller, Darrell Long, and Lee Ward\\
Source of Support           \>: \> ASCR\\
Award Amount                \>: \> Approximately \$1,500,000\\
Period Covered              \>: \> 10/2015-9/2018\\
Commitment                  \>: \> 0.1 FTE\\
Description                 \>: \> Develop local and global mechanisms for managing non-volatile memories integrated \\
 \> \>into nodes of supercomputers.\\
Overlap                     \>: \> None\\
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Modeling Impacts of Resilience Architectures for Extreme-Scale Storage Systems\\
Principal Investigator      \>: \> Remzi Arpaci-Dusseau, Matthew L. Curry, and Nathan DeBardeleben\\
Source of Support           \>: \> ASCR\\
Award Amount                \>: \> Approximately \$1,500,000\\
Period Covered              \>: \> 10/2015-9/2018\\
Commitment                  \>: \> 0.25 FTE\\
Description                 \>: \> Project to model multi-level resilience mechanisms for large-scale storage systems\\
Overlap                     \>: \> None\\
\end{tabbing}

\end{document}