### 2015-07-08 coverpage.tex

%------------------------------------------------------------------------------
%                 P r o p o s a  l   C o v e r   P a g e
%------------------------------------------------------------------------------
\thispagestyle{empty}

\begin{center}
\textbf{\Large \baseproposaltitle}
\end{center}

\begin{center}
%\includegraphics[width=1in]{LOGO.gif}
\end{center}

\medskip

%% Application/Institution
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Applicant Institution:}} \= Oak Ridge National Laboratory \\
\>  \>1 Bethel Valley Road\\
\> \> Oak Ridge Tennessee 37831-6057\\
\end{tabbing}

\medskip

%% Lead PI name, telephone number, email:
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Lead Principal Investigator:}} \=
\vrule height0.4pt depth 0pt width 2.5in \\
\> \> Scott Klasky,  Distinguished Scientist \\
\> \> Oak Ridge National Laboratory \\
\> \> Computer Science and Mathematics Division \\
\> \> Oak Ridge, Tennessee 37831-6057 \\
\> \> (T) 865-241-9980 (Fax) 865-241-4811 \\
\> \> Email: klasky@ornl.gov
\end{tabbing}

\medskip

%% Adminisrative Point of Contact name, telephone number, email:
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Administrative Point of Contact:}} \=
\vrule height0.4pt depth 0pt width 2.5in \\
\> \> Barney Maccabe \\
\> \> Oak Ridge National Laboratory \\
\> \> Computer Science and Mathematics Division \\
\> \> Oak Ridge, TN 37831-6057 \\
\> \> (T) 865-241-6504 \\
\> \> Email: maccabe@ornl.gov
\end{tabbing}

\bigskip

\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Program Announcement Number:}} \= LAB 15-1338\\
\> \textbf{PAMS Tracking No.:}			    \> 0000219877\\
\> \textbf{DOE/Office of Science Program Office:} \> Office of Advanced Scientific Computing Research\\
\>									    \> Storage Systems and Input/Output for Extreme Scale Science\\
\> \textbf{DOE/Office of Science Program Office}  \> \\
\> \makebox[64mm][r]{\textbf{Technical Contact:}} \>  Lucy Nowell\\
\\
\> \textbf{Research Area:} \>Scalable Storage Software Infrastructure

\end{tabbing}

\newpage