### 2015-07-08 participant-table.tex

% Generated by bin/config-participant-table on Thu Jun 18 15:07:03 EDT 2015
\newlength{\longestpientry}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}}}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}}}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}}}
\begin{tabularx}{\textwidth}{l>{\raggedright\arraybackslash}X}
\emph{Institution} & \makebox[\longestpientry][l]{\emph{Principal Investigator (Email)}} \quad \emph{Additional Senior Personnel}
\\
\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Gary Liu, Kimmy Mu\\
\textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\
\textbf{\rutgers} & \makebox[\longestpientry][l]{\textbf{Manish Parashar} (\emph{parashar@rutgers.edu})}\hspace*{1em}\\
\textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}
\end{tabularx}