### 2015-07-09 other-support/ward-cp.tex

\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{cite}
\usepackage{url}
\usepackage{times}
\usepackage{graphicx}
\usepackage[sectionbib]{bibunits}
\pagestyle{empty}

% Set page geometry
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

\begin{center}
{\Large Current and Pending Support for H. Lee Ward}
\end{center}

Mr. Ward is a principle member of the technical staff in the Center for
Computing Research group at Sandia National Laboratories (SNL). The following
table identifies current and pending support. This funding provides for effort,
travel, and resources. The effort levels will be adjusted if pending proposals
are funded, and as current projects finish.

\section*{Current Support}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Sirocco\\
Principal Investigator      \>: \> Lee Ward\\
Source of Support           \>: \> DOE NNSA/Advanced Simulation and Computing\\
Period Covered              \>: \> 10/2014-9/2015\\
Commitment                  \>: \> 0.85 FTE\\
Description                 \>: \> Development activities for Sirocco, a novel storage systemfor Exascale computers.\\
Overlap                     \>: \> None\\
\end{tabbing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Pending Proposals (in addition to this)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Pending Proposals (in addition to this)}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Exascale Storage Authorization and Data-Location Research\\
Principal Investigator      \>: \> Anthony Skjellum\\
Source of Support           \>: \> DOE Office of Science\\
Period Covered              \>: \> 10/2015-9/2018\\
Commitment                  \>: \> 0.058 FTE\\
Description                 \>: \> Design, consulting, mentoring in support of research using Sirocco as a vehicle.\\
Overlap                     \>: \> None\\
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Managing Node-Based Non-Volatile Memories for HPC Clusters\\
Principal Investigator      \>: \> Ethan Miller\\
Source of Support           \>: \> DOE Office of Science\\
Period Covered              \>: \> 10/2015-9/2018\\
Commitment                  \>: \> 0.25 FTE\\
Description                 \>: \> Design and development activities for Sirocco in support of research using Sirocco as a vehicle.\\
Overlap                     \>: \> None\\
\end{tabbing}

\end{document}