### 2015-07-09 coverpage.tex

%------------------------------------------------------------------------------
%                 P r o p o s a  l   C o v e r   P a g e
%------------------------------------------------------------------------------
\thispagestyle{empty}

\begin{center}
\textbf{\Large \baseproposaltitle}
\end{center}

\begin{center}
%\includegraphics[width=1in]{LOGO.gif}
\end{center}

\medskip

%% Application/Institution
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{Applicant/Institution:} \= Oak Ridge National Laboratory \\
\>  \>1 Bethel Valley Road\\
\> \> Oak Ridge Tennessee 37831-6057\\
\end{tabbing}

%\medskip

%% Postal Address
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{Postal Address:} \= P. O. Box 2008 \\
%\> \> Oak Ridge National Laboratory \\
%\> \> Computer Science and Mathematics Division \\
\> \> Oak Ridge, Tennessee 37831-6057 \\
%\> \> (T) 865-241-9980 (Fax) 865-241-4811 \\
%\> \> Email: klasky@ornl.gov
\end{tabbing}
%\vrule height0.4pt depth 0pt width 2.5in \\

%\medskip

%% Adminisrative Point of Contact name, telephone number, email:
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{Administrative POC:} \= Barney Maccabe \\
%\> \> Oak Ridge National Laboratory \\
%\> \> Computer Science and Mathematics Division \\
%\> \> Oak Ridge, TN 37831-6057 \\
\> \> Phone: 865-241-6504 \\
\> \> Email: maccabe@ornl.gov
\end{tabbing}
%\vrule height0.4pt depth 0pt width 2.5in \\

%\medskip

%% Lead PI name, telephone number, email:
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{Lead PI:} \= Scott Klasky, Oak Ridge National Laboratory \\
%\> \> Oak Ridge National Laboratory \\
%\> \> Computer Science and Mathematics Division \\
%\> \> Oak Ridge, Tennessee 37831-6057 \\
%\> \> (T) 865-241-9980 (Fax) 865-241-4811 \\
%\> \> Email: klasky@ornl.gov
\end{tabbing}
%\vrule height0.4pt depth 0pt width 2.5in \\

%\medskip

\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{DOE National Lab Announcement No.:} \= \textbf{LAB 15-1338}\\
\\
\> {PAMS Letter of Intent Tracking Number:}			    \> \textbf{0000219877}\\
\\
\> {DOE/Office of Science Program Office:} \> \textbf{Office of Advanced Scientific Computing}\\
\>									    \> \textbf{Research Storage Systems and Input/Output}\\
\>									    \> \textbf{(SSIO) for Extreme Scale Science}\\
\\
\> {DOE/Office of Science}  \> \\
\> \makebox[64mm][r]{Program Office Technical Contact:} \>  \textbf{Lucy Nowell}\\
\\
\> {Research Area:} \>\textbf{Scalable Storage Software Infrastructure}

\end{tabbing}

\newpage