### 2015-07-10 narrative/proposed-work.tex
fixed a few thing in the proposed work.

\section{Proposed Research and Methods}
\label{subsec:challenges}
\label{sec:proposed}

%\begin{enumerate}
%\item Re-factoring of the data
%  \begin{itemize}
%  \item What kind of reorganization would work when?
%  \item Would it take too much time and resources + memory and how can we
%    manage the cost of the reorganization
%  \item Will the users loose too much value from the data during refactoring
%    and how can we control/bound this loss of value
%  \end{itemize}
%\item Quality of Service
%  \begin{itemize}
%  \item Can we make any guarantees in a scalable decentralized system?
%  \item What quality of service metrics can be make guarantees on in a scalable
%    way?
%  \item What kind of error rates will we have in our estimation and what error
%    rates are acceptable?
%  \item How much will the qos scheduler, admission control, manager cost in
%    terms of resources and time
%  \item How do we make the scheduler scalable with the size of the system?
%  \end{itemize}
%\item Discovery and naming
%  \begin{itemize}
%  \item Which data block and what quality/type do we respond to a query
%    with?
%  \item How will we scale a fully distributed storage system's metadata
%    management
%  \item Can we accelerate the discovery process with the addition of an
%    application/user level metadata store that tracks where data was
%    originally written and how was it structured?
%  \item Can we provide time estimates for read/write calls in the presence of
%    an unbounded discovery system? If not how do we bound the discovery time?
%  \item Time cost of discovery vs time cost of actually reading the data
%    (latency vs bandwidth essentially)
%  \end{itemize}
%\item Migration
%  \begin{itemize}
%  \item Can we do scalable migration without making the discovery process
%    unbounded
%  \item What are the parameters and input for migration
%  \item Can we use the additional application level knowledge of data to purge
%    portions of the data from storage without making the data too much less
%    valuable?
%  \item With so many objects and no centralized directory how do we know what
%    is actually in the storage to be purged?
%  \item How do purge in a way that doesn't require a central authority,
%    doesn't interfere with other I/O and isn't bottlenecked by some sort of
%    global consistency
%  \end{itemize}
%\end{enumerate}

%Our research into the storage system and middle ware layers must address
%many of these challenges including
% \begin{enumerate}
% \item How do we describe the user intentions at the API layer and have this
%   communicated down to the middle-ware and storage layers?
% \item How do we allow users to define their user defined compression
%   techniques and have this data read back from the storage system layers?
%   Where do we execute the decompression techniques 
% \item How do we evaluate the tradeoffs at runtime to guide data placement,
%   including the movement of data across the network, across the different
%   storage tiers? (data movement and quality tradeoff)
% \item Can we use different forms of learning techniques as daemons on the
%   system to perform data migration from one storage tier to another? For
%   example, if we see that a user is looking at one time slice after another
%   for a certain object in their dataset which is in the slower tiers of the
%   hierarchy, do we automatically propane up the data which has not been
%   requested in the hope that this will be requested data? (prediction and
%   prefetching, user defined compression on movement)
% \item Can we understand the true need of campaign storage and understand the
%   possible impact of the campaign storage if we run a hadoop-like
%   file-system instead of a lustre/gpfs file system? (additional layers)
% \item Can we limit the amount of data duplication? Data space is going to be
%   very constrained on the exascale system so we must ensure that minimal
%   copies are made of data, and when data is duplicated, they are removed
%   through a garbage collection routine. (space management)
% \item Can we build a model to give us the time estimates during the reading
%   and writing phases, so that rules can be in place from the user
%   perspective to make adaptive decisions. For example, if the file system
%   says that reading will take 3 months, users can they place in rules to
%   then read in a subset of data, and they can understand which data can be
%   read quickly and which pieces will take more time. (estimation)
% \item Can we understand how to place code in the system which will allow
%   data-regeneration to take place. (regeneration)
% \item How does the storage system do a better job in managing request from
%   all of the users on a LCF than today? We realize that today users who have
%   a better middlware system can often lock other users from getting high
%   performance when they are running. If we have the concept of currency
%   which is eventually used in the same extent as node-hours, then users will
%   have to be able to think about how much storage and how much bandwidth
%   they can choose. One question that needs to be understood is if there are
%   times when the system sees that there are very few storage system
%   resources being used, then the lucky users can get the bandwidth cheaper
%   than at times of heavy usage. How can we enforce this? (Fairness)
% \item How do we manage all of the metdata not only from the principle
%   objects, but from the sub-objects? How well will this scale when we have
%   users who can potentially create billions of objects from their
%   simulation? (scalability and discovery)

%\paragraph{Deliverables and artifacts}
%
%\begin{enumerate}
%\item Add plug-in architecture to Sirocco to support selective data compression
%\item Add specific tier destinations command to Sirocco (assuming that the caching mechansism cannot achieve the desired effects).
%\item Develop profiling system to determine how a data set is used during preparation runs prior to a capability run to determine how to optimize data placement for capability run analytics.
%
%\end{enumerate}
%
%{\bf \color{red}we need to condense these challenges into a few basic principles - 4 at most - SAK}

We propose to overcome these challenges by pioneering a new {\bf
knowledge-centric approach} to a dynamic storage system and I/O layer which
takes into account user prioritization and utility to optimize the SSIO layer.
The approach is based on two underlying principles:

\underline{Principle 1:} {\bf A knowledge-centric system design} that allows user knowledge  to
define data policies. Today SSIO layers are written in a
stove-pipe fashion, and quite often do not allow optimizations to take place.
We propose to re-design the layers in a highly integrated fashion where users
will place their intentions into the system and actions will statically and dynamically
take place to optimize for both the system and for individual
requests.

\underline{Principle 2}: {\bf Predictable performance and quality of data in the SSIO layers} needs to be
established so science can be done on the exascale systems in a more efficient manner. Without
predictable performance, not only can the runs be slowed down because of contentions \cite{liu_hotstorage}, 
but also it affects key science decisions, e.g., how much data reduction should be performed.
It is clear that if a run is expected to be slow, an aggresive data reduction needs to be performed
so that science can be done in a timely fashion, which has been reflected in the XGC run (\S~\ref{sec:introduction})
that the PI recently participated.

We aim to alleviate the need for the ``magic'' and ``tricks''
that are currently required to optimize application I/O performance
on today's file systems. To accomplish this, we will provide a systematic
approach for describing intentions and other knowledge from {\it user}
space, as well as allow performance estimations and guarantees 
from the underlyingree storage.
%For example, ADIOS contains most of the ``best practices'' for I/O optimizations for
%   each file system, and as users place more knowledge into ADIOS, ADIOS
%   can make optimizations.
By capturing user intentions and acting upon them in the middleware, we
free the user from polluting application code with system specific
optimizations which may provide better performance in the short term,
but are detrimental to the long-term maintainability of applications, and
are more sensitive to changes to system configuration. We have successfully
employed this strategy of separation of concerns in ADIOS, and believe that
it will become increasingly important because saving all data may not be possible, and
users want the ability to describe and prioritize different chunks of data. 
Our research is aimed to allow users to provide simple descriptions of their data,
which further allows the middleware and storage to optimize application I/O performance.

One of the underlying principles of our approach is that we recognize that there are a small set
of important I/O patterns to optimize for~\cite{lofstead2011six,polte2009and,tian2011edo,tian2012system},
and our representative applications contain all of these I/O patterns. Throughout our project
we will continuously focus on creating research artifacts which can optimize these patterns on current  and
exascale systems.  One important pattern is Checkpoint/Restarts (CR) and we will focus on optimizing this pattern,
 and also using many of the important research techniques for using NVRAM and neighbor's NVRAM for storing and retrieving
the CR data. The migration and purging of this will be optimized by our SSIO layer using the utility function of the CR data so that this can  be stored for short periods of time, and then purged.

%\underline{Principle 3:} {\bf Users goals define data policies} In a data-centric SSIO system. 
%We introduce data-driven SSIO optimizations by
%enabling users to define data manipulations (re-organizations) at a high level
%and then we express policy goals by associating intentions (hints) with data;
%e.g. ``this data at this level should be accessible within 100 seconds in the
%next three months.'' These high level specifications of desired results will be
%used in our system for data migration and eviction policies.

\input{narrative/data-description} % includes data-desc, apis, and qos-app.

% \subsection{Application Interface}

%\input{narrative/data-desc}

\input{narrative/data-reduction} % includes data-reorg and the api specs move to the above section -
%\input{narrative/data-reorg}

%\input{narrative/qos-app}
\input{narrative/placement}
% \input{narrative/migration}
% \input{narrative/eviction-metrics}

%\input{narrative/policies}
%\input{narrative/apis}

% \subsection{Storage System}

%\input{narrative/resource-mgmt}
\input{narrative/qos}
%\input{narrative/discovery}

\input{narrative/naming-service}
% \input{narrative/time-estimation}
\input{narrative/infrastructure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End: