### 2015-07-10 cover-collab.tex
supplement for collaboration

\vfill
\begin{center}
\textbf{\huge Cover Page Supplement for Collaborations}
\end{center}

\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Model content for table, really included below
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \begin{center}
%% \textbf{\large Collaborating Institutions}
%% \medskip
%% \begin{tabular}{lll}
%% \hline\hline
%%   & \emph{Institution} & \emph{Principal Investigator} \\
%% \hline
%% \textbf{Lead PI} & \textbf{Oak Ridge National Laboratory} 
%%   & \textbf{Sudharshan Vazhkudai}\\
%% Co-PI & Lawrence Berkeley National Laboratory & Ramakrishnan Lavanya \\
%% \hline\hline
%% \end{tabular}
%% \end{center}

\input{cover-collab-table}

\vfill
\begin{center}
\textbf{\large Leadership Structure}
\medskip
\end{center}

This is a collaborative proposal between 
Oak Ridge National Laboratory, and
Sandia National Laboratories, and
Rutgers University, and
University of California, Santa-Cruz.

%% Provide a statement explaining the leadership structure of the
%% collaboration.

\vfill
\begin{center}
\textbf{\large Facilities, Equipment, and Resources}
\medskip
\end{center}

This proposal brings together the expertiese and facilities from two
national laboratories, Oak Ridge National Laboratory and Sandia National
Laboratories, and two universities, Rutgers University and University of
California, Santa-Cruz. 
 In addition, by
sharing experiese and personnel costs, it represents
a collaboration with other large programs within the Office of
Science, including on data management and exascale monitoring, as well as
ongoing development of the Sirocco file-system. 
The proposed work will leverage existing facilities at all three
sites, including leadership machines at ORNL and SNL, and also utilize
storage oriented cluster systems at all locations. The PIs will also combine
their experience with production middleware and storage system software
development and their collaboration with leading application scientists to
better guide the assumptions underlying this effort.

%% Include a description of each collaborating institution’s facilities,
%% equipment, and resources that will be made available to the
%% collaborative group.

\vfill
\begin{center}
\textbf{\large Summary Budget Information}
\medskip
\begin{tabular}{| l | l | l | r | r | r | r |} \hline
Role & Name & Institution & Year 1 (\$K) & Year 2 (\$K) &   Year 3 (\$K) & Total (\$K) \\
\hline
Lead-PI & Scott Klasky & ORNL & {\$}570 &  {\$}570 &  {\$}570 &   {\$}1,710\\
Co-PI & Jay Lofstead & Sandia  &  {\$}350 &  {\$}350 &  {\$}350 &  {\$}1,050 \\
Co-PI & Carlos  Maltzahn & UCSC  &  {\$}160 &  {\$}160 &  {\$}160 &  {\$}480 \\
Co-PI & Manish Parashar & Rutgers &  {\$}170 &  {\$}170 &  {\$}170 &  {\$}510  \\
\hline
{\bf Totals}  & & & {\bf  {\$}1,250} & {\bf  {\$}1,250} & {\bf  {\$}1,250} & {\bf  {\$}3,750}\\
\hline
\end{tabular}
\end{center}

\noindent
{Period of Performance: October 1, 2015 - September 31, 2018}\\
\noindent
\vfill