### 2015-08-12 abstract.tex
Modified the abstract following Lucy's directive to shorten to half a page

% Special formatting for the abstract
\titleformat{\chapter}[block]
%{\filcenter\normalfont\huge\bfseries}{}{0pt}{\huge}
{\filcenter\normalfont\LARGE\bfseries}{}{0pt}{\LARGE}
%{\filcenter\normalfont\Large\bfseries}{}{0pt}{\Large}
%{\filcenter\normalfont\Large\bfseries}{}{0pt}{\large}
\titlespacing*{\chapter}{0pt}{-0.35in}{0pt} %%% * kill first para indentation

\chapter*{\MakeUppercase{\proposaltitle}}
\phantomsection
%\addstarredchapter{Abstract}
%%\adjustmtc
\markboth{Abstract}{}
\label{sec:abstract}

% Restore standard formatting for chapter titles
\titleformat{\chapter}[block]
{\normalfont\Huge\bfseries}{\thechapter.}{10pt}{\Huge}
\titlespacing*{\chapter}{0pt}{-0.25in}{0pt} %%% * kill first para indentation

%% \ifthenelse{\equal{\proposalcontent}{abstract}}{ % Standalone abstract-only
%%    \typeout{output-controls: empty pagestyle for standalone abstract}
%%    \thispagestyle{empty}
%% }{}

% Vertical spacing is careful controlled in order to maximize space
% available while maintaining a reasonable experience.

\vspace{-\abovedisplayskip}
\medskip

% Lead PI slug is centered internally
\textbf{\input{leadpi-slug}}

%\vspace{-\belowdisplayskip}
\vspace{-\abovedisplayskip}
\smallskip

\begin{center}
\input{participant-table}
\end{center}

\vspace{-\belowdisplayskip}
\vspace{-\abovedisplayskip}
\medskip

\begin{center}
\textbf{ABSTRACT}
\end{center}

\vspace{-\belowdisplayskip}

% Even with the space manipulation above, we need to cheat a little
\enlargethispage{2\baselineskip}

Scientific discovery at the exascale will not be possible without significant
new research into the management, storage and retrieval over the full lifespan
of the extreme amounts of data that will be produced. Our thesis is that by
adding application level knowledge about data to guide the actions of the
storage system we obtain substantial benefits in the organization, storage, and
access to extreme scale data, resulting in improved productivity for
computational science. We will demonstrate novel techniques to facilitate
efficient mapping of data objects from the user space onto multiple storage
tiers, and enable application-guided data reductions and transformations to
address capacity and bandwidth bottlenecks.  Our goal is to address the
associated Input/Output (I/O) and storage challenges in the context of current
and emerging storage landscapes, and expedite insights into mission critical
scientific processes.  We aim to reduce the time to insight, not just for a
single application, but for the entire workload in a multi-user environment,
where storage is shared across users. We shall achieve this by allowing
selectable data quality, through trading in accuracy and error in order to meet
the time or resource constraint. We will explore beyond checkpoint/restart I/O,
and will address the challenges posed by key data access patterns in the
knowledge gathering process. Ultimately, we will take the knowledge from the
storage system to provide vital feedback to the middleware so that the best
possible decisions can be autonomically made between the user intentions and
the available system resources.  We will test our prototypes on current and
future Department of Energy (DOE) system with many of today's cutting edge
applications to ensure our techniques can be used within our framework on
current and next generation systems.

\noindent

\vfill