### 2015-10-21 proposal/acronym.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

%%%%% Acronym definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \SVN $Id$

% Avoid printing the chapter heading if we don't want the acronyms printed
\ifthenelse{\boolean{printacro}}{
   \chapter{Glossary of Acronyms}
   \typeout{DEBUGGING: In acronym.tex chapter heading}
}{\newpage}

\typeout{DEBUGGING: In acronym.tex}

%%% Definitions of acronyms: \acro{id}[short name]{full name}
%%% In text, reference with \ac{id}
\begin{acronym}

% Autogenerated institutional acronyms with all-lower and all-upper versions
\input{acronym-inst}

% Autogenerated acronyms for the institutional PIs
\input{acronym-pis}

%%% Definitions of acronyms: \acro{id}[short name]{full name}

\acro{AACE}[AACE]{Architecture-Aware Compilation Environment}
\acro{ACTS}[ACTS]{Advanced Computational Software}
\acro{ALCF}[ALCF]{Argonne Leadership Computing Facility}
\acro{AM}[AM]{applied math}
\acro{APGAS}[APGAS]{asynchronous partitioned global address space}
\acro{API}[API]{application program interface}
\acro{ASCI}[ASCI]{Accelerated Strategic Computing Initiative}
\acro{ASCR}[ASCR]{Office of Advanced Scientific Computing Research}
\acro{ASC}[ASC]{Advanced Simulation and Computing}
\acro{AST}[AST]{abstract syntax tree}
\acro{BES}[BES]{Office of Basic Energy Sciences}
\acro{BOFs}[BOFs]{``Birds of a Feather'' sessions}
\acro{CASC}[CASC]{Center for Applied Scientific Computing}
\acro{CASL}[CASL]{Consortium for Advanced Simulation of Light Water Reactors}
\acro{CBHPC}[CBHPC]{Component-Based High-Performance Computing}
\acro{CCA}[CCA]{Common Component Architecture}
\acro{CCTTSS}[CCTTSS]{Center for Component Technology for Terascale Simulation Software}
\acro{CESAR}[CESAR]{Center for Exascale Simulation of Advanced Reactors}
\acro{CESM}[CESM]{Community Earth System Model}
\acro{CGM}[CGM]{Common Geometry Module}
\acro{CSE}[CSE]{computational science and engineering}
\acro{CSP}[CSP]{communicating sequential processes}
\acro{CS}[CS]{computer science}
\acro{DAMSEL}[DAMSEL]{Data Model Storage Library for Exascale Science}
\acro{DARPA}[DARPA]{Defense Advanced Research Projects Agency}
\acro{DFT}[DFT]{density functional theory}
\acro{DOE}[DOE]{Dept.~of Energy}
\acro{DSL}[DSL]{domain-specific language} % Cite with \acp to get plural
\acro{DVI}[DVI]{differential variational inequality}
\acro{ECDC}[ECDC]{Exascale Co-Design Consortium}
\acro{eDSL}[eDSL]{embedded domain-specific language}
\acro{EMSL}[EMSL]{Environmental Molecular Sciences Laboratory}
\acro{ESMF}[ESMF]{Earth System Modeling Framework}
\acro{ExMatEx}[ExMatEx]{Exascale Co-Design Center for Materials in Extreme Environments}
\acro{ExaCT}[ExaCT]{Exascale Simulation of Combustion in Turbulence}
\acro{FACETS}[FACETS]{Framework Application for Core-Edge Transport Simulations}
\acro{FASTMath}[FASTMath]{Frameworks, Algorithms, and Scalable Technologies for Mathematics Institute}
\acro{FES}[FES]{Office of Fusion Energy Sciences}
\acro{FLOPS}[FLOPS]{Floating Point Operations Per Second}
\acro{FNAL}[Fermilab]{Fermi National Accelerator Laboratory}
\acro{FOA}[FOA]{Funding Opportunity Announcement}
\acro{FSP}[FSP]{Fusion Simulation Program}
\acro{GPU}[GPU]{graphical processing unit}
\acro{GUI}[GUI]{graphical user interface}
\acro{HPC}[HPC]{high-performance computing}
\acro{IDE}[IDE]{interactive development environment}
\acro{INCITE}[INCITE]{Innovative and Novel Computational Impact on Theory and Experiment}
\acro{IPM}[IPM]{Integrated Performance Monitoring}
\acro{IPS}[IPS]{Integrated Plasma Simulator}
\acro{IR}[IR]{intermediate representation}
\acro{ISA}[ISA]{instruction set architecture}
\acro{ITAPS}[ITAPS]{Interoperable Tools for Advance Petascale Simulation}
\acro{IT}[IT]{Information Technology}
\acro{LCF}[LCF]{Leadership Computing Facility}
\acro{MCMD}[MCMD]{Multiple-Component Multiple-Data}
\acro{MC}[MC]{Monte Carlo}
\acro{MDD}[MDD]{model-driven design}
\acro{MD}[MD]{molecular dynamics}
\acro{MPI}[MPI]{Message Passing Interface}
\acro{MSCF}[MCSF]{Molecular Science Computing Facility}
\acro{MS}[M\&S]{modeling and simulation}
\acro{NDA}[NDA]{non-disclosure agreement}
\acro{NERSC}[NERSC]{National Energy Research Scientific Computing Center}
\acro{NNSA}[NNSA]{National Nuclear Security Administration}
\acro{NSF}[NSF]{National Science Foundation}
\acro{NUMA}[NUMA]{non-uniform memory access}
\acro{OASCR}[ASCR]{Office of Advanced Scientific Computing Research}
\acro{OLCF}[OLCF]{Oak Ridge Leadership Computing Facility}
\acro{OO}[OO]{object-oriented}
\acro{OS}[OS]{operating system}
\acro{PCIe}[PCIe]{Peripheral Component Interconnect Express}
\acro{PGAS}[PGAS]{partitioned global address space}
\acro{QA}[QA]{quality assurance}
\acro{RD}[R\&D]{research and development}
\acro{RDMA}[RDMA]{remote direct memory access}
\acro{SC}[SC]{Office of Science}
\acro{SDMAV}[SDMAV]{the anticipated SciDAC Institute in Scientific Data Management, Analysis, and Visualization}
\acro{SE}[SE]{software engineering}
\acro{SIMD}[SIMD]{single-instruction multiple-data}
\acro{SPMD}[SPMD]{single-program multiple-data}
\acro{SQA}[SQA]{software quality assurance}
\acro{SST}[SST]{Structured Simulation Toolkit}
\acro{SUPER}[SUPER]{Institute for Sustained Performance, Energy, and Resilience}\acro{QUEST}[QUEST]{Quantification of Uncertainty in Extreme Scale Computations}\acro{FASTMath}[FASTMath]{Frameworks, Algorithms, and Scalable Technologies for Mathematics}
\acro{SciDAC}[SciDAC]{Scientific Discovery through Advanced Computing}
\acro{TASCS}[TASCS]{Center for Technology for Advanced Scientific Component Software}
\acro{TCE}[TCE]{Tensor Contraction Engine}
\acro{TDD}[TDD]{test-driven development}
\acro{UHPC}[UHPC]{Ubiquitious High-Performance Computing}
\acro{UI}[UI]{user interface}
\acro{UML}[UML]{Unified Modeling Language}
\acro{UM}[UM]{University of Maryland}
\acro{UQ}[UQ]{uncertainty quantification}
\acro{VCS}[VCS]{version control system}
\acro{VV}[V\&V]{verification and validation}

\acro{LULESH}[LULESH]{Livermore Unstructured Lagrangian Explicit Shock Hydrodynamics}
\acro{SSCA}[SSCA]{Scalable Synthetic Compact Application}
\acro{DoD}[DoD]{Dept.~of Defense}
\acro{HPCMP}[HPCMP]{High Performance Computing Modernization Program}

\acro{DVFS}[DVFS]{dynamic voltage and frequency scaling}
\acro{TMR}[TMR]{triple modular redundancy}
\acro{ABFT}[ABFT]{algorithm-based fault tolerance}
\end{acronym}