### 2015-12-23 papers/2016-HotStorage/paper.tex
added template for Hot Storage paper

\documentclass{sig-alt-gov2}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{graphicx}
%\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}

\newcommand{\DDT}{D\textsuperscript{2}T~}
\newcommand{\DDTns}{D\textsuperscript{2}T}

\hyphenation{sub-trans-ac-tion}

\addtolength{\parskip}{-0.08in}

% in order for balance columns to work, it has to be before begin document...
%\balancecolumns

\begin{document}

%\conferenceinfo{PDSW'15,} {November 17, 2015, Austin, Texas}
%\CopyrightYear{2015}
%\crdata{978-1-4503-1103-8/11/11}
%\clubpenalty=10000
%\widowpenalty = 10000

\title{A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing}

\numberofauthors{1}
\author{
%\alignauthor {Jay Lofstead\footnotemark[1], Ivo Jimenez\footnotemark[2],Carlos Maltzahn\footnotemark[2], Quincey Koziol\footnotemark[3], John Bent\footnotemark[4], Eric Barton\footnotemark[5]}\\
\alignauthor {Jay Lofstead\footnotemark[1], Carlos Maltzahn\footnotemark[2], Scott Klasky\footnotemark[3], Hasan Abbasi\footnotemark[3], Qing Liu\footnotemark[3]\\
Matthew Curry\footnotemark[1], Mark Ainsworth\footnotemark[5], Manish Parashar\footnotemark[4]}\\
\footnotemark[1]~~Sandia National Laboratories gflofst@sandia.gov\\
\footnotemark[2]~~University of California, Santa Cruz carlosm@soe.ucsc.edu\\
\footnotemark[3]~~Oak Ridge National Laboratory\\
\footnotemark[4]~~Rutgers University\\
\footnotemark[5]~~Brown University\\
}
\maketitle

\begin{abstract}

The longest existing parallel I/O APIs, HDF5, PnetCDF, and NetCDF, are focused
on treating the entire output from a paralell application as a single unit.
These libraries work to gather the data from all processes and arrange it such
that it is organized much like as if it were written from a single process.
This approach, in particular the two-phase collective I/O with data sieving
portion, has proven difficult to scale for 3-D domain decompositions at scale.
ADIOS shifted the focus from treating the entire application output as a single
entity to treating the output from each writer as a self-contained entity. This
eliminated the need for the problematic two-phase collective I/O and has
demonstrated good scalability well into the petascale era. Another shift is
required to address the needs of exascale systems. New system I/O bandwidth is
not keeping pace with the compute acceleration leading to accute bottlenecks.
While ``burst buffer'' technology will help address the performance gap, it
does not address the problem completely. This paper describes a new approach
to not just an I/O library, but also the underlying storage infrastructure to
address the needs of exascacle applications cognizant of projected exascale
platform characteristics and in the context of current industry trends.

\end{abstract}

%\category{D.4}{Software}{Operating Systems}
%\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]

%\terms{Design, Performance}

\section{Introduction}

%\begin{figure}[htbp]
%\vspace{-0.10in}
%\centering
%\includegraphics[width=\columnwidth]{images/arch-mapping}
%\vspace{-0.20in}
%\caption{Architecture and Component Mapping}
%\label{fig:arch-mapping}
%\vspace{-0.15in}
%\end{figure}

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array.
Section~\ref{sec:storage}
discusses the challenges and opportunities for deploying scalable storage
infrastructure.
Section~\ref{sec:qos} 
discusses the quality of service features proposed for this project.
Next is an exploration of the metadata challenges required to support such a
system in Section~\ref{sec:metadata}.
An initial prototype demonstration of the functionality is presented in
Section~\ref{sec:evaluation}.
The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

\section{Related Work}
\label{sec:related}

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in parallel
file systems.
Ceph's unique striping approach uses pseudo-random numbers with a
known seed eliminating the need for the metadata service to track where each
stripe in a parallel file is placed.
PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
such as a single process opening a file and sharing the handle.
It has been
commercialized in recent years as OrangeFS.
Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
clusters offering scalable performance and fine-grained end-user and
programmatic control over how data is placed in the storage system.
GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good
performance for scaling parallel IO tasks and is used extensively by its owner,
IBM.
Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive
striping system that detects the need for additional stripes for performance
and adjusts the file layout as necessary.

Other file systems, like GoogleFS~\cite{ghemawat:googlefs} and
HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
computing and cannot be compared directly.
The primary difference between
distributed and parallel file systems is the ability of the file system to
store and retrieve data simultaneously from multiple clients, in parallel, and
treat the resulting collection of pieces as a single object.  Distributed file
systems rely on a single client creating a file, but distributing the set of
files across a wide array of storage devices.
The other, popular distributed
file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
decades for enterprise file systems.
These other file systems are mainly of
interest in the context of the ACG features of FFSIO and will be discussed more
in Section~\ref{sec:acg}.

\section{End-User API Layer}
\label{sec:end-user}

\section{Storage Challenges and Opportunities}
\label{sec:storage}

\section{Quality of Service Features}
\label{sec:qos}

\section{Metadata Challenges}
\label{sec:metadata}

\section{Demonstration}
\label{sec:evaluation}

This stack has an early prototype implementation intended to test concepts
rather than performance and scalability. It has focused on examining the
interaction of the different APIs for each layer to flesh out any detailed
requirements or concerns that may have been missed in the conceptualization of
this IO stack. To demonstrate the viability of the IO stack described in this
paper, we show some very early performance results from the untuned prototype.

\section{Conclusions}
\label{sec:conclusion}

\section{Acknowledgements}

\includegraphics[scale=0.07]{logos/doe_logo}
\includegraphics[scale=0.30]{logos/snl_logo}
\includegraphics[scale=0.35]{logos/nnsa_logo}
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

\bibliographystyle{abbrv}
\bibliography{paper}

\vfill\eject

\end{document}