### 2016-02-05 papers/2016-Supercomputing/paper.tex
initial work on the Supercomputing paper

\documentclass{sig-alt-gov2}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{graphicx}
%\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}

\hyphenation{sub-trans-ac-tion}

\addtolength{\parskip}{-0.08in}

% in order for balance columns to work, it has to be before begin document...
%\balancecolumns

\begin{document}

%\conferenceinfo{PDSW'15,} {November 17, 2015, Austin, Texas}
%\CopyrightYear{2015}
%\crdata{978-1-4503-1103-8/11/11}
%\clubpenalty=10000
%\widowpenalty = 10000

\title{Restructuring Data for Exascale Scientific Productivity}

\numberofauthors{1}
\author{
\alignauthor {
Jay Lofstead\footnotemark[1]~~,
Hasan Abbasi\footnotemark[3]~~,
Mark Ainsworth\footnotemark[6]~\footnotemark[3]~~,
Jong Choi\footnotemark[3]~~,
Matthew Curry\footnotemark[1]~~,\\
Scott Klasky\footnotemark[3]~\footnotemark[4]~\footnotemark[7]~~,
Tashin Kurc\footnotemark[9]~~,
Qing Liu\footnotemark[3]~~,
Carlos Maltzahn\footnotemark[2]~~,
Manish Parashar\footnotemark[5]~~,\\
Norbert Podhorzski\footnotemark[3]~~,
Feiyi Wang\footnotemark[3]~~,
Matthew Wolf\footnotemark[7]~~,
C. S. Chang\footnotemark[8]~~,
Michael Churchill\footnotemark[8]~~,
Stephane Ethier\footnotemark[8]~~
}\\
\footnotemark[1]~~Sandia National Laboratories,
\footnotemark[2]~~University of California Santa Cruz,\\
\footnotemark[3]~~Oak Ridge National Laboratory,
\footnotemark[4]~~U. Tenn. Knoxville,
\footnotemark[5]~~Rutgers University,\\
\footnotemark[6]~~Brown University,
\footnotemark[7]~~Georgia Tech,
\footnotemark[8]~~PPPL,
\footnotemark[9]~~Stony Brook
}
\maketitle

\begin{abstract}

As the exascale computing age emerges, data related issues are becoming
critical factors that determine how and where we do computing. Popular
approaches used by traditional I/O solution and storage libraries become
increasingly bottlenecked due to their assumptions about data movement,
re-organization, and storage. While, new technologies, such as ``burst
buffers'', can help address some of the short-term performance issues, it is
essential that we reexamine the underlying storage and I/O infrastructure to
effectively support requirements and challenges at exascale and beyond.

In this paper we present
- different data needs different techniques
- write overhead for the different refactoring/auditing approaches
- derived quantities in viz after precision based refactoring and differences
  from full fidelity data sets.
- linear auditing approaches for data refactoring and the effects on data
  sizes, io performance, and viz quality. Do every point, but split at 24 bits
and 40 bits.

\end{abstract}

%\category{D.4}{Software}{Operating Systems}
%\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]

%\terms{Design, Performance}

\section{Introduction}

%\begin{figure}[htbp]
%\vspace{-0.10in}
%\centering
%\includegraphics[width=\columnwidth]{images/arch-mapping}
%\vspace{-0.20in}
%\caption{Architecture and Component Mapping}
%\label{fig:arch-mapping}
%\vspace{-0.15in}
%\end{figure}

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array.
Section~\ref{sec:iof}
briefly discusses the motivation and proposal for the IO forwarding layer.
Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
functionality it offers.
Section~\ref{sec:daos} discusses how the DAOS
layer functions.
The VOSD layer is discussed in Section~\ref{sec:vosd}.
Next is an exploration of cross-cutting features like transactions and
metadata management in Section~\ref{sec:broader}.
A demonstration of the functionality is presented in
Section~\ref{sec:evaluation}.
The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

\section{Related Work}
\label{sec:related}

Many projects over the last couple of decades have sought to address some
challenging aspect of parallel file system design. The recent rise of ``Big
Data'' applications with different characteristic IO patterns have somewhat
complicated the picture. Extreme scale machines will be expected to handle both
the traditional simulation-related workloads as well as applications more
squarely in the Big Data arena. This will require some adjustments to the
underlying system for good performance for both scenarios.

The major previous work is really limited to full file systems rather than the
mountain of file system refinements made over the years. A selection of these
other file systems and some features that make it relatively unique are
described below.

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in parallel
file systems.
Ceph's unique striping approach uses pseudo-random numbers with a
known seed eliminating the need for the metadata service to track where each
stripe in a parallel file is placed.
PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
such as a single process opening a file and sharing the handle.
It has been
commercialized in recent years as OrangeFS.
Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
clusters offering scalable performance and fine-grained end-user and
programmatic control over how data is placed in the storage system.
GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good
performance for scaling parallel IO tasks and is used extensively by its owner,
IBM.
Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive
striping system that detects the need for additional stripes for performance
and adjusts the file layout as necessary.

Other file systems, like GoogleFS~\cite{ghemawat:googlefs} and
HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
computing and cannot be compared directly.
The primary difference between
distributed and parallel file systems is the ability of the file system to
store and retrieve data simultaneously from multiple clients, in parallel, and
treat the resulting collection of pieces as a single object.  Distributed file
systems rely on a single client creating a file, but distributing the set of
files across a wide array of storage devices.
The other, popular distributed
file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
decades for enterprise file systems.
These other file systems are mainly of
interest in the context of the ACG features of FFSIO and will be discussed more
in Section~\ref{sec:acg}.

\section{Demonstration}
\label{sec:evaluation}

\section{Conclusions}
\label{sec:conclusion}

The Fast Forward Storage and IO Stack project has designed a good first pass at
addressing the requirements for an extreme scale data storage mechanism.  By
preferring a high level user API like HDF5 rather than using the POSIX
interface, more advanced functionality can be incorporated with less end-user
impact. The introduction of the IOD layer with buffering will absorb the
difference between the compute node IO demands and the available bandwidth in
the storage array. With DAOS supporting translating the container and object
model to the underlying storage options, different storage technologies can be
deployed over time.

With the overall stack design a prototype implementation complete, refinements,
such as fault detection and recovery, can be designed and tested.  These and
other activities for phase 2 will ultimately generate what is likely to be the
next generation storage stack for extreme scale platforms.

\section{Acknowledgements}

Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

\bibliographystyle{abbrv}
\bibliography{paper}

\vfill\eject

\end{document}