### 2016-03-11 papers/2016-HotStorage/paper.tex
finished paper.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes,url}
\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf SIRIUS: A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing}

\author{
Jay Lofstead\footnotemark[1]~~,
Hasan Abbasi\footnotemark[3]~~,
Mark Ainsworth\footnotemark[6]~\footnotemark[3]~~,
Jong Choi\footnotemark[3]~~,
Matthew Curry\footnotemark[1]~~,\\
Scott Klasky\footnotemark[3]~\footnotemark[4]~\footnotemark[7]~~,
Tahsin Kurc\footnotemark[9]~~,
Qing Liu\footnotemark[3]~~,
Jeremy Logan\footnotemark[4]~~,
Carlos Maltzahn\footnotemark[2]~~,
Manish Parashar\footnotemark[5]~~,\\
Norbert Podhorzski\footnotemark[3]~~,
Feiyi Wang\footnotemark[3]~~,
Matthew Wolf\footnotemark[7]~~,
C. S. Chang\footnotemark[8]~~,
Michael Churchill\footnotemark[8]~~,
Stephane Ethier\footnotemark[8]~~
\\
\footnotemark[1]~~Sandia National Laboratories,
\footnotemark[2]~~University of California Santa Cruz,\\
\footnotemark[3]~~Oak Ridge National Laboratory,
\footnotemark[4]~~U. Tenn. Knoxville,
\footnotemark[5]~~Rutgers University,\\
\footnotemark[6]~~Brown University,
\footnotemark[7]~~Georgia Tech,
\footnotemark[8]~~PPPL,
\footnotemark[9]~~Stony Brook
}

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}

\subsection*{Abstract}
The longest existing parallel I/O APIs, HDF5, PnetCDF, and NetCDF, are focused
on treating the entire output from a parallel application as a single unit.
These libraries work to gather the data from all processes and arrange it such
that it is organized much like as if it were written from a single process.
This approach, in particular the two-phase collective I/O with data sieving
portion, has proven difficult to scale for 3-D domain decompositions at scale.
ADIOS shifted the focus from treating the entire application output as a single
entity to treating the output from each writer as a self-contained entity. This
eliminated the need for the problematic two-phase collective I/O and has
demonstrated good scalability well into the petascale era. Another shift is
required to address the needs of exascale systems. New supercomputing system
I/O bandwidth is not keeping pace with the compute acceleration leading to
accute bottlenecks.  While ``burst buffer'' technology will help address the
performance gap, it does not address the problem completely. This paper
describes a new approach to not just an I/O library, but also the underlying
storage infrastructure to address the needs of exascacle applications cognizant
of projected exascale platform characteristics and in the context of current
industry trends.

\section{Introduction}

High performance computing simulations have driven tightly coupled platform
scaling for decades. The output from these simulations are analyzed using
regular tools and techniques optimized for this environment.  The first
generation of standardized APIs for scientific computing IO focused on serial
output from a single process to a distributed or local file system.  These
systems still exist in the form of serial
HDF5~\cite{chilan:2006:hdf5-performance} and the end-of-life NetCDF
3.x~\cite{netcdf3}. The key feature of these APIs is the logical data model
that eliminates any parallel data decomoposition artifacts from the stored
data.  This offered consistent performance and regular data access patterns.

The second generation is really an evolution of the first with HDF-5 adding
parallel capabilities and NetCDF bifurcating into
NetCDF4~\cite{rew:2011:netcdf-guide} and PnetCDF~\cite{torrellas:2003:pnetcdf}
with PnetCDF almost completely maintaining the data model and NetCDF4 shifting
to use HDF5 underneath. The other improvement is incorporating knowledge of an
underlying parallel file system to support large files written in unision from
many processes.

The third generation is a reaction to the second based on recognizing that
maintaining the logical data model in physical form became inordinately
expensive~\cite{lofstead:2011:six-degrees} and offered a performance penalty in
most cases when running at scale. Libraries like
ADIOS~\cite{lofstead:2009:adaptable} shifted from writing complete variables
from distributed processes as a single entity to treating all data from a
single source to be treated as a unit. This eliminated the data reorganization
at the cost of more complicated reading patterns. In spite of this extra
complication, the performance is generally
better~\cite{lofstead:2011:six-degrees}.

SIRIUS represents the next generation. New platforms are introducting
additional storage layers to accelerate IO operations and to compensate for
plateauing storage array performance. At the same time, data sizes continue
to grow. The fourth generation IO API will have to both address the IO API
and middleware layer as well as the deeper, more complex storage hierarchy.
SIRIUS considers the local portion of each globally distributed variable as a
management entity with data placement decisions across the entire storage
hierarhy rather than on a single tier.  Decisions on what data to place where
is driven by the data contents itself according to end-user specified
priorities and system characteristics. For example, if temperature and pressure
are nearly always retrieved together, organizing the data to keep just these
two arrays together without any interleaving from other variables or scalars
will accelerate later reading. If these variables could also be ``processed''
such that the important data features are maintained in high resolution while
less interesting regions are stored in highly compressed--even using lossy
compression--forms, the amount of data written can be greatly increased. While
accelerating writing is important. This approach also supports accelerating
reading.

This paper presents an overview of the efforts to build such a fourth
generation storage and IO system. A discussion of the platform features,
storage challenges, quality of service features, and metadata challenges are
all discussed briefly.

%\begin{figure}[htbp]
%\vspace{-0.10in}
%\centering
%\includegraphics[width=\columnwidth]{images/arch-mapping}
%\vspace{-0.20in}
%\caption{Architecture and Component Mapping}
%\label{fig:arch-mapping}
%\vspace{-0.15in}
%\end{figure}

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:features}
discusses the differentiating platform features SIRIUS will provide.
Section~\ref{sec:storage}
discusses the challenges and opportunities for deploying scalable storage
infrastructure.
Section~\ref{sec:qos}
discusses the quality of service features proposed for this project.
Next is an exploration of the metadata challenges required to support such a
system in Section~\ref{sec:metadata}.
An initial prototype demonstration of the functionality is presented in
Section~\ref{sec:evaluation}.
The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

\section{Related Work}
\label{sec:related}

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in
parallel file systems.  Ceph's unique striping approach uses pseudo-random
numbers with a known seed eliminating the need for the metadata service to
track where each stripe in a parallel file is placed.  PVFS~\cite{carns:pvfs}
offers optimizations to reduce metadata server load, such as a single process
opening a file and sharing the handle.  It has been commercialized in recent
years as OrangeFS.  Lustre~\cite{braam:lustre-arch} has become the de facto
standard on most major clusters offering scalable performance and fine-grained
end-user and programmatic control over how data is placed in the storage
system.  GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing
good performance for scaling parallel IO tasks and is used extensively by its
owner, IBM.  Panasas~\cite{panasas:architecture} seeks to offer a dynamically
adaptive striping system that detects the need for additional stripes for
performance and adjusts the file layout as necessary.

Other file systems, like GoogleFS~\cite{ghemawat:googlefs} and
HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
computing and cannot be compared directly.  The primary difference between
distributed and parallel file systems is the ability of the file system to
store and retrieve data simultaneously from multiple clients, in parallel, and
treat the resulting collection of pieces as a single object.  Distributed file
systems rely on a single client creating a file, but distributing the set of
files across a wide array of storage devices.  The other, popular distributed
file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
decades for enterprise file systems.

\section{Platform Features}
\label{sec:features}

The platform has two key features that drive the rest of the design. First is
the data refactoring/compression approach. Second are quality of service
estimates.

\subsection{Refactoring}
Refactoring may be a uniform operation across an entire variable. For example,
splitting a double into three parts--two byes, four bytes, and two bytes,
yields a two-byte value that is the majority of the data precision. The next
four bytes would be most of the rest of the data precision. The last two bytes
are the remainder of the precision. For quick overviews, just reading the
two-byte portion is sufficient to get a rough idea of data contents. If
particular data looks interesting, more and more data can be read to improve
the precision. This can easily be applied uniformly by each process without
coordinating. From a storage hierarchy perspective, the primary value portion
could be written to high performance, low capacity storage while the rest is
written to slower storage. That would reduce the data quantity written to NVM,
for example, by 75\% for a 2/6 split enabling far more output to be written to
this fast tier with the rest spooled asynchronously.

A similar approach that maintains full data quality is to use a data stride to
store only a data subset. For example, by selecting every fourth element in
three dimensions yields a $\frac{1}{64}$ sized data set to store in fast
storage--a mere 1.5\% of the original size. The rest can be spooled
asynchronously. Combining byte-splits and striding can reduce the data sizes
even more. Doing a 3/5 split and an every fourth element yields 0.9\% of the
original size.

A more complex, but interesting approach is to offer an {\em Auditor} that
performs a similar, but lower precision computation that greatly reduces
computation intensity with only a minor loss of precision over short spatial
and temporal domains. As the simulation progresses, the auditor drift would
need to be reset to bring it back into alignment. The much smaller, reduced
precision data could be used instead of the full precision in many
circumstances.  For an example auditor, consider a previous generation physics
model. While it is not as accurate as the state-of-the-art-model, it is very
close, particularly over short temporal and spatial ranges. Other types of
auditors are possible and are under investigation. The key idea with the
auditor is that the reduced quality result can be calculated much more quickly
than the full precision and will have much smaller data sizes. As a side
benefit, the auditor can both detect soft errors in the caluclation by the
drift from the high precision physics diverging too quickly. It can also be
used to create a statistically representative data set for the high precision
physics model. Calculating the auditor value and writing its output for
checkpoint restart or initial analysis exploration can save time and leverage
otherwise idle excess compute resources.

A more dynamic lossy approach would be to apply a wavelet
compression~\cite{klappenecker:1995:wavelet} approach that generates different
compressed data sizes at a fixed error rate depending on the data entropy.
These are still uniform operators, but the resulting data can be of radically
different sizes. For this approach, some or all of the lower precision data
can be written to the fast tier while full precision can be spooled
asynchrnously to the slower tier(s).

A middle ground approach between these refactoring approaches is to use a
non-uniform operator across a variable. Instead of, for example, splitting on
byte boundaries to make multiple values, a function that identifies data
features through a local analytics function could identify ``interesting'' data
that should be maintained in full precision while other areas can be more
highly compressed. Such an operator would decompose the data from a process
further than a single process or node depending on the data features. This
approach may balance precision better than wavelet compression, but at a higher
analytics cost. SIRIUS would put data areas with interesting features on the
fast tier while the rest could be spooled to the slow tier asynchronously.

We plan to investigate all of these approaches as different applications will
have different techniques that work best. Our initial tests have demonstrated
the potential for the auditor approach in an isolated environment. We have also
built a prototype ADIOS transport method for the byte-split apporach. Any
performant storage system should offer the flexibility to incorporate
application specific data refactoring techniques and sufficient metadata
support for an arbitrary client to access the data successfully.

\subsection{Quality of Service}
Quality of service deadlines are important for driving how much time is spent
performing IO. At the same time, data quality must be maintained at a
sufficient level for the resulting science to be possible.  The system must
offer a way for an end-user to negotiate with the system for trading off
between time spent in IO and the amount of data written or read.  Part of that
tradeoff will be increased or reduced data quality, depending on the system.
For example, when data is written, a few different forms may be output given
sufficient space and time. One may be every fourth element across a regular
mesh reduces the data size to $1/64^{th}$ the full data set.  Further dividing
that using byte-splitting can reduce it by another 75\%. For quick overviews to
search data sets, this reduced form can be sufficient driving which data sets
should be investigated in more detail. The key driving factor is time. The
applicaiton scientist can read 0.4\% of the full data volume with a similar
time reduction. SIRIUS will offer a negotiation process to read any data.

\begin{table}[tbp]
\centering
\caption{Reading Options for Variable ``A''}
\label{tab:results}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\# & Size & Time & Time Err & Refactoring & Data Err\\
\hline
1 & $(\frac{1}{4})^3(\frac{3}{8})$A & 10s & $\pm$ 3s & stride, byte-split & 98\% \\
\hline
2 & $(\frac{3}{4})^3(\frac{3}{8})$A & 90s & $\pm$ 30s & stride, byte-split & 58\% \\
\hline
3 & $(\frac{1}{4})^3(\frac{5}{8})$A & 16s & $\pm$ 5s & stride, byte-split & n/a \\
\hline
4 & $(\frac{3}{4})^3(\frac{5}{8})$A & 120s & $\pm$ 50s & stride, byte-split & n/a \\
\hline
\hline
5 & $(\frac{1}{4})^3$A & 1200s & $\pm$ 30s & stride & 98\% \\
\hline
6 & $(\frac{3}{4})^3$A & 2400s & $\pm$ 90s & stride & 58\% \\
\hline
\hline
7 & $(\frac{3}{8})$A & 1350s & $\pm$ 120s & byte-split & 0.02\% \\
\hline
8 & $(\frac{5}{8})$A & 2250s & $\pm$ 120s & byte-split & n/a \\
\hline
\hline
9 & A & 36s & $\pm$ 6s & wavelet & 1\% \\
\hline
\hline
10 & A & 3600s & $\pm$ 600s & none & 0\% \\
\hline
\end{tabular}
\end{table}

The byte splitting approach error rate is quite small. By splitting at byte
boundaries rather than converting to a lower precision 4-byte float, we
preserve the full exponent range and keep the first 11 bits of mantissa
precision. This leaves a $2^{-12}$ maximum error magnitude for each value--the
last 40 bits represent less precision together than the immediately preceeding
bit all by itself.

To read these various data organizations, read operations will shift to a
two-step process. First, an list of options will be requested that includes
both and error estiates. A second request will retrieve a particular data
portion from the provided list. Consider Table~\ref{tab:results}.  First, a
user will request a particular variable and timesetp. SIRIUS will return a
table similar to Table~\ref{tab:results}. For a given variable ``A'',
the end-user can then select the data quality and approximate reading time for
the operation. In this case, there are 5 options for obtaining the whole data
set (1/2/3/4, 5/6, 7/8, 9, 10). The first four entries represent a spatial and
byte split, 5/6 represent just strided writing, 7/8 represent just a
byte-split, 9 represents a wavelet compressed version, and 10 is the full
precision data.  The user would specify to read a particular data instance that
would proceed normally.  Quality of Service soft guarantees specified as part
of the request will result in SIRIUS allocating soft bandwidth reservations to
meet the request. While some of the error rates seem to render the data
unusable, consider a 3000x3000x3000 simulation domain. That would yield a 27
gigapixel image should all of the data be used. Visualization will already
radically reduce the data sizes to make usable visualizations that work on a
standard display. A 99\% reduction would yield a 270 megapixel image--still
quite large.

\section{Storage Challenges and Opportunities}
\label{sec:storage}

The supercomputing platforms being deployed today are all incorporating a
Non-volatile memory (NVM) layer as a fast cache between compute and the storage
array. The assumption is that nearly all IO will be performed against this
layer to hide the much greater latency to the storage array. The challenge is
the NVM capacity is expensive limiting the amount deployed. The unspoken
limitation is that this fast layer still must drain to the relatively slow
storage array. This performance mismatch will limit capacity and performance
just due to the drain and data stage-in operations.

A more effective way to use these resources is to only store the most
important data in the fast tier and store lower precision representations of
the rest of the data in the slower tiers. This lowers pressures on the NVM
layer increasing performance overall by greatly reducing the drain/stage-in
operations and reducing the data volumes stored on a per application basis.
As part of the data refactoring, SIRIUS will use the entire storage hierarchy
to store different portions of each variable as illustrated in
Table~\ref{tab:results}. For example, for the 1/2/3/4 setup, 4 different tiers
may be used to store this single variable.

\section{Quality of Service Features}
\label{sec:qos}

SIRIUS offers a novel feature related to managing IO time as a first class
characteristic. Rather than being at the mercy of the storage array
performance, an end-user can describe the local data to the storage system and
receive an estimate on how long that data will take to write. If that is
acceptable within a margin of error, the API can write. If it is unacceptable,
the end-user can make decisions about skipping the output or maybe spending a
little time with a strong data refactoring to reduce the data size.  On
reading, a user will select from a list of available options based on the time
and data quality requirements.

Key to making QoS work is to maintain performance metadata to inform the IO
time estimates and error bounds returned to the user. Without hard
reservations, no system can offer strict performance guarantees. Instead, by
monitoring system behavior, SIRIUS will be able to offer reasonable time
estimates given data placement and the requesting source(s).

\section{Metadata Challenges}
\label{sec:metadata}

The metadata system has a daunting challenge. Traditional metadata systems for
both distributed and parallel file systems only have to deal with a single name
space and storage tier. While hierarichal storage
management~\cite{blaze:1992:hsm} systems offer a way to show a single name
space, the reality is that the tier closest to the user is a cache for the
other tiers with data migrating as necessary to support user interaction.
SIRIUS is working in a fundamentally different way. Each tier is treated as a
first class citizen with direct client access possible. Where data is actually
retrieved from is where it is actually stored rather than forcing data
migration and caching. Systems like Sirocco~\cite{curry:2015:sirocco} offer a
way to support this model directly and are a fundamental part of our system.

The additional challege SIRIUS metadata services face is incorporating
user-specified attributes including data refactoring functions. Since an
arbitrary user must be able to access any data written for which they have
permissions. Having data locked behind data compression that has changed over
time may result in effectively encrypting data and losing the decryption key.

We also seek to offer the ability to also store multiple versions of the same
data using different refactoring techniques. This will offer additional options
for higher precision OR higher performance reading.

\section{Demonstration}
\label{sec:evaluation}

Our initial tests are on the Sith cluster at Oak Ridge. The system contains 40
compute nodes. Each compute node contains four 2.3 GHz 8 core AMD Opteron
processors, and 64 GB of memory. The system is configured with a 86 TB Lustre
file system for scratch space.

We use the XGC1~\cite{ku:2008:xgc1} fusion simulation as our application. We
run on 5 nodes with 160 processes total. XGC1 is configured to use 819,200,000
particles evenly spread across the processes. Performance numbers are
presented in Table~\ref{tab:performance}.

\begin{table}[tbp]
\centering
\caption{Performance For 3/5 byte Split}
\label{tab:performance}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
Test & Time (s)\\
\hline
Write full data & 10.3 \\
\hline
Byte split & 2.5 \\
\hline
write 3-byte portion & 3.4 \\
\hline
write 5-byte portion & 9.3 \\
\hline
\hline
Read 3-byte portion & 10.3 \\
\hline
Reassemble after read & 27.9 \\
\hline
Read full data & 34.2 \\
\hline
\end{tabular}
\end{table}

Evaluating the error when using just the 3-byte portion compared to the full
data set yields an $L^2$ norm of 72028.2 and $L_{\inf}$ of 0.00109242.

Writing the 3-byte portion synchronously to fast storage and spooling the
5-byte portion out to slower storage asynchronously will preserve the full data
fidelity while improving both writing and reading time even when including the
disassemble/reassemble times. The introduced errors when using just the 3-byte
portion, as discussed above, is quite small.

\section{Conclusions}
\label{sec:conclusion}

Overall, the SIRIUS storage system is ushering in a new generation of both IO
APIs and storage system managed together. It incorporates additional metadata
to help accelerate data access and ultimately time to insight. Our initial
efforts show promise that such a system is both possible and can offer
attractive performance to better use the supercomputing platform. By treating
all storage tiers as first-class resources accessed independently, SIRIUS can
maximize available IO bandwidth. By refactoring data to extract out important
parts into separate data set slices, data writing and reading time can be
greatly improved while still maintaining full data fidelity when necessary.
The quality of service monitoring will inform the performance estimates giving
a more predictably responsive system accelerating time to insight.

%\subsection{How to Build Your Paper}
%
%You have to run {\tt latex} once to prepare your references for
%munging.  Then run {\tt bibtex} to build your bibliography metadata.
%Then run {\tt latex} twice to ensure all references have been resolved.
%If your source file is called {\tt usenixTemplate.tex} and your {\tt
%  bibtex} file is called {\tt usenixTemplate.bib}, here's what you do:
%{\tt \small
%\begin{verbatim}
%latex usenixTemplate
%bibtex usenixTemplate
%latex usenixTemplate
%latex usenixTemplate
%\end{verbatim}
%}
%

\section{Acknowledgments}

This project is funded by the following research grants from the Advanced
Scientific Research Office at the Department of Energy, under contracts SDAV:
ERKJ200/KJ0403000, RSVP: ERKJU60/KJ0402000, and Sirius: ERKJ311/KJ0402000.
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

{\footnotesize \bibliographystyle{acm}
\bibliography{paper}}

%\theendnotes

\end{document}