### 2016-04-08 papers/2016-Supercomputing/details.tex
Add interpolation details. Add self as author.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

Modern experiments and numerical simulations produce datasets so large that their storage and retrieval strain the capabilities of storage hardware.
Slow writing of the data is problematic on shared machines where memory is periodically purged and transferred to new users; slow reading delays analysis of the data after the experiment is over.
Compression offers a solution to this problem.
A crude but commonly used compression technique is decimation, wherein only every \(\stride\)th datum of a sequence is stored.
In practice, the decimation factor \(\stride\), also known as the stride, is often chosen to be \(10\).
The remaining data is simply discarded.
Clearly, decimation trades a reduction in the size of the dataset for data loss.
The loss can be mitigated by generating approximations to the discarded data from the retained data.
A simple, natural technique for doing so is linear interpolation.
Let \(\Data\) be the original dataset.
To generate an approximation \(\Datum[approx]{j}\) to a discarded datum \(\Datum{j}\), we find stored values \(\Datum{m\stride}\) and \(\Datum{m\stride+\stride}\) that straddle \(\Datum{j}\) (\(m\stride<j<m\stride+\stride\)) and take a weighted average:
\begin{equation*}
\Datum{j} \approx \Datum[approx]{j} = (1-\tfrac{j-m\stride}{\stride})\Datum{m\stride}+\tfrac{m\stride+\stride-j}{\stride}\Datum{m\stride+\stride}
.
\end{equation*}
Observe that decimation paired with interpolation, illustrated in Fig.~\ref{fig:InterpolationReconstructionOfDiscardedValues}, gives a compression ratio of \(\stride\) and provides a full set of values at all original timesteps.
\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth]{images/interpolation/illustration_of_procedure}
\caption{Illustration of decimation procedure for partitioning of data in the case \(\stride=5\).}
\label{fig:InterpolationReconstructionOfDiscardedValues}
\end{figure}
This procedure produces a good compression ratio, and it is computationally attractive: for a given \(j\) only two stored values need to be retrieved in order to reconstruct a value for \(\Datum{j}\).
However, as noted above, there is a direct tradeoff between the compression ratio and the data loss.
Nonetheless, despite these obvious shortcomings, the above approach or variants thereof is in widespread use as a technique for compression large datasets, particularly in the context of computational simulation.
In our implementation, the data is partitioned into the \emph{top level data} \(\DataDecimated\), consisting of the decimated data, and the \emph{bottom level data} \(\InterpolationErrors\), consisting of the differences, or \emph{deltas}, between the original data and the approximated data.

Typically, the deltas resulting from decimation paired with interpolation are simply discarded.
The tacit assumption behind this approach is that the approximated dataset \(\Data[approx]\) is a sufficiently good approximation to the full dataset \(\Data\).
The issue of the accuracy of this process has received little or no attention, but the following result is proved in \cite{AinsworthWhitneyKlasky}:
\begin{equation}
\max_{j\in\NN}\lvert\InterpolationError{j}\rvert \leq \min(
2\MaximumDifferenceGlobal{0},
\tfrac{1}{2}\stride\MaximumDifferenceGlobal{1},
\tfrac{1}{8}\stride^{2}\MaximumDifferenceGlobal{2}
)
\label{eqn:InterpolationMagnitudeOfDeltasBound}
\end{equation}
where \(\MaximumDifferenceGlobal{0}=\max_{j\in\NN}\lvert\Datum{j}\rvert\), \(\MaximumDifferenceGlobal{1}=\max_{j\in\NN}\lvert\Datum{j}-\Datum{j+1}\rvert\), and \(\MaximumDifferenceGlobal{2}=\max_{j\in\NN}\lvert\Datum{j}-2\Datum{j+1}+\Datum{j+2}\rvert\).
These bounds are illustrated for a particular dataset in Fig.~\ref{fig:InterpolationDeltaMagnitudeExample}.
Notice that when \(\stride\) is smaller, the accuracy is high, meaning that the top level data produces an acceptable surrogate for the full original dataset.
However, for larger strides (or, equally well, for higher compression ratios), the accuracy becomes unacceptable.
The bottom level data, containing the deltas, enables lossless reconstruction of the data when the accuracy of the approximations is unacceptable for a particular purpose.
\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth]{images/interpolation/delta_sizes}
\caption{%
Illustration of the bounds in Eq.~\eqref{eqn:InterpolationMagnitudeOfDeltasBound} for a range of strides in the case \(\Datum{j}=\sin(2^{-5}j)\) for \(j=0,\ldots,\lfloor2^{6}\pi\rfloor\).
Observe the transition between different cases of Eq.~\eqref{eqn:InterpolationMagnitudeOfDeltasBound} at \(\stride\approx2^{7}\).%
}
\label{fig:InterpolationDeltaMagnitudeExample}
\end{figure}

One might question whether this partitioning procedure provides any overall compression of the original data.
The amount by which a dataset can be compressed is quantified by its Shannon entropy.
One advantage of the decimation--interpolation process is that for reasonably smooth data the entropy of the deltas \(\InterpolationErrors\) will be smaller than that of the original data stream \(\Data\).
Consequently, the deltas are more amenable to compression using an entropy encoder than were the original data.
The following result is proved in \cite{AinsworthWhitneyKlasky}.

\begin{theorem}
\label{thm:InterpolationDifferentialEntropyBounds}
The average differential entropy per delta in \(\InterpolationErrors\) is bounded by the minimum of
\begin{enumerate}[(1)]
\item \(1+\log_{2}\MaximumDifferenceGlobal{0}+\tfrac{1}{\stride-1}\log_{2}\big[\tfrac{1}{6}(\stride+1)(\stride+2)\big]\)
\item \(1+\log_{2}\MaximumDifferenceGlobal{1}\)
\item \(1+\log_{2}\MaximumDifferenceGlobal{2}-\tfrac{1}{\stride-1}\log_{2}\stride\)
\end{enumerate}
\end{theorem}

These entropy bounds can be translated to bounds on the expected compression ratio achieved by the combination of decimation, interpolation, and compression of the deltas.
See \cite{AinsworthWhitneyKlasky} for details and Fig.~\ref{fig:InterpolationCompressionRatioExample} for an illustration.

\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth]{images/interpolation/compression_ratios}
\caption{%
Illustration of the compression ratio bounds derived in \cite{AinsworthWhitneyKlasky} from Theorem~\ref{thm:InterpolationDifferentialEntropyBounds} for a range of strides in the case \(\Datum{j}=2^{4}(3+\sin(2^{-15}j))\) for \(j=0,\ldots,\lfloor2^{16}\pi\rfloor\).%
}
\label{fig:InterpolationCompressionRatioExample}
\end{figure}